---
title: "Untitled"
author: "First Last"
date: YYYY-MM-DD
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```


# Introduction

The following RMarkdown file uses files from Donors Choose and performs preliminary sentiment analysis. The sentiment analysis was aided by the code and explanations provided in  https://www.tidytextmining.com/. 

# Get Started

```{r}
# Libraries

library(dplyr)
library(stringr)
library(tidytext)
library(ggplot2)
library(tidyverse)
library(cowplot)

# Open the datasets

train <- read.csv("~/Dropbox/DonorsChoose/train.csv")
test <- read.csv("~/Dropbox/DonorsChoose/test.csv")
resources <- read.csv("~/Dropbox/DonorsChoose/resources.csv")

```

# Data exploration on the title

## Make data into a tibble and find out how many words are in the title

```{r}
# First, we want to select project id, title name, and if the project was approved or not

id_title <- c(1, 16, 9)

train_text <- train[,id_title]
train_text[,1] <- as.character(train_text[,1])
train_text[,2] <- as.numeric(train_text[,2])
train_text[,3] <- as.character(train_text[,3])

train_text <- as.tibble(train_text)

tidy_books <- train_text %>% unnest_tokens(word, project_title)

# Find out how many words the title is

freq_table <- count(tidy_books, id) 

title_word_count_by_project <- left_join(freq_table, train_text[,1:2], by = c("id"))

ggplot(title_word_count_by_project, aes(x = factor(project_is_approved), y = n)) + geom_boxplot()

ggplot(title_word_count_by_project, aes(x = n, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill")
```

## Take out common words ("stop words")

```{r}
# Take out stop (common) words

tidy_books <- tidy_books %>%
  anti_join(stop_words)

freq_table <- count(tidy_books, id) 

title_word_count_by_project <- left_join(freq_table, train_text[,1:2], by = c("id"))

ggplot(title_word_count_by_project, aes(x = factor(project_is_approved), y = n)) + geom_boxplot()

title_n <- ggplot(title_word_count_by_project, aes(x = n, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill") + ggtitle("Number of words by approval status")

# Get a sense of which words are the most common

tidy_books %>%
  count(word, sort = TRUE) 


tidy_books %>%
  count(word, sort = TRUE) %>%
  filter(n > 2000) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip()

```

## Sentiment analysis

### AFINN

```{r}
janeaustensentiment <- tidy_books %>% inner_join(get_sentiments("afinn"))

summary(janeaustensentiment)

janeaustensentiment[1:10,]

# Is there a correlation between the average score and whether or not it gets accepted?

check_corr_titles <- aggregate(janeaustensentiment$score, by = list(janeaustensentiment$id), FUN = mean)
colnames(check_corr_titles) <- c("id", "score")

check_corr_titles2 <- left_join(check_corr_titles, train_text[,1:2], by = c("id"))


ggplot(check_corr_titles2, aes(x = factor(project_is_approved), y = score)) + geom_boxplot()

title_a <- ggplot(check_corr_titles2,aes(x = score, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill") + ggtitle("Avg. AFINN sentiment score by approval status")

# Is there a relationship between the most positive word and whether it gets approved?

check_corr_titles <- aggregate(janeaustensentiment$score, by = list(janeaustensentiment$id), FUN = max)
colnames(check_corr_titles) <- c("id", "score")

check_corr_titles2 <- left_join(check_corr_titles, train_text[,1:2], by = c("id"))


ggplot(check_corr_titles2, aes(x = factor(project_is_approved), y = score)) + geom_boxplot()

title_b <- ggplot(check_corr_titles2,aes(x = score, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill") + ggtitle("Max. AFINN sentiment score by approval status")

# Is there a relationship between the most negative word and whether it gets approved?

check_corr_titles <- aggregate(janeaustensentiment$score, by = list(janeaustensentiment$id), FUN = min)
colnames(check_corr_titles) <- c("id", "score")

check_corr_titles2 <- left_join(check_corr_titles, train_text[,1:2], by = c("id"))


ggplot(check_corr_titles2, aes(x = factor(project_is_approved), y = score)) + geom_boxplot()

title_c <- ggplot(check_corr_titles2,aes(x = score, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill") + ggtitle("Min. AFINN sentiment score by approval status")

# Is there a relationship between the most common value and whether it gets approved?

check_corr_titles <- aggregate(janeaustensentiment$score, by = list(janeaustensentiment$id), FUN = median)
colnames(check_corr_titles) <- c("id", "score")

check_corr_titles2 <- left_join(check_corr_titles, train_text[,1:2], by = c("id"))


ggplot(check_corr_titles2, aes(x = factor(project_is_approved), y = score)) + geom_boxplot()

title_d <- ggplot(check_corr_titles2,aes(x = score, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill") + ggtitle("Median AFINN sentiment score by approval status")

```


### BING

```{r}
# Sentiment analysis - bing

janeaustensentiment <- tidy_books %>% inner_join(get_sentiments("bing"))

janeaustensentiment[1:10,]

# Is there a connection between the average score and whether or not it gets accepted?

length(which(janeaustensentiment$project_is_approved == 1 & janeaustensentiment$sentiment == "positive"))
length(which(janeaustensentiment$project_is_approved == 0 & janeaustensentiment$sentiment == "positive"))

length(which(janeaustensentiment$project_is_approved == 1 & janeaustensentiment$sentiment == "negative"))
length(which(janeaustensentiment$project_is_approved == 0 & janeaustensentiment$sentiment == "negative"))


bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(id) %>%
  summarize(words = n())

sentiment_ratio <- tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(id, project_is_approved) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("id")) %>%
  mutate(ratio = negativewords/words) %>%
  top_n(1) %>%
  ungroup()

ggplot(sentiment_ratio, aes(x = factor(project_is_approved), y = ratio)) + geom_boxplot()


```

### NRC

```{r}
janeaustensentiment <- tidy_books %>% inner_join(get_sentiments("nrc"))

janeaustensentiment[1:20,]

```

# Data exploration on essay 1

## Make data into a tibble and find out how many words are in essay 1

```{r}
# First, we want to select project id, title name, and if the project was approved or not

id_title <- c(1, 16, 10)

train_text <- train[,id_title]
train_text[,1] <- as.character(train_text[,1])
train_text[,2] <- as.numeric(train_text[,2])
train_text[,3] <- as.character(train_text[,3])

train_text <- as.tibble(train_text)

tidy_books <- train_text %>% unnest_tokens(word, project_essay_1)

# Find how many words in essay1

freq_table <- count(tidy_books, id) 

title_word_count_by_project <- left_join(freq_table, train_text[,1:2], by = c("id"))

ggplot(title_word_count_by_project, aes(x = factor(project_is_approved), y = n)) + geom_boxplot()

ggplot(title_word_count_by_project, aes(x = n, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill")

```

## Take out common words ("stop words")

```{r}
# Take out stop (common) words

tidy_books <- tidy_books %>%
  anti_join(stop_words)

freq_table <- count(tidy_books, id) 

title_word_count_by_project <- left_join(freq_table, train_text[,1:2], by = c("id"))

ggplot(title_word_count_by_project, aes(x = factor(project_is_approved), y = n)) + geom_boxplot()

essay1_n <- ggplot(title_word_count_by_project, aes(x = n, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill")

# Get a sense of which words are the most common

tidy_books %>%
  count(word, sort = TRUE) 


tidy_books %>%
  count(word, sort = TRUE) %>%
  filter(n > 20000) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip()

```

## Sentiment analysis

### AFINN

```{r}
janeaustensentiment <- tidy_books %>% inner_join(get_sentiments("afinn"))

summary(janeaustensentiment)

janeaustensentiment[1:10,]

# Is there a relationship between the average score and whether or not it gets accepted?

check_corr_titles <- aggregate(janeaustensentiment$score, by = list(janeaustensentiment$id), FUN = mean)
colnames(check_corr_titles) <- c("id", "score")

check_corr_titles2 <- left_join(check_corr_titles, train_text[,1:2], by = c("id"))

ggplot(check_corr_titles2, aes(x = factor(project_is_approved), y = score)) + geom_boxplot()

essay1_a <- ggplot(check_corr_titles2,aes(x = score, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill") 


# Is there a relationship between the most positive word and whether it gets approved?

check_corr_titles <- aggregate(janeaustensentiment$score, by = list(janeaustensentiment$id), FUN = max)
colnames(check_corr_titles) <- c("id", "score")

check_corr_titles2 <- left_join(check_corr_titles, train_text[,1:2], by = c("id"))


ggplot(check_corr_titles2, aes(x = factor(project_is_approved), y = score)) + geom_boxplot()

essay1_b <- ggplot(check_corr_titles2,aes(x = score, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill")

# Is there a relationship between the most negative word and whether it gets approved?

check_corr_titles <- aggregate(janeaustensentiment$score, by = list(janeaustensentiment$id), FUN = min)
colnames(check_corr_titles) <- c("id", "score")

check_corr_titles2 <- left_join(check_corr_titles, train_text[,1:2], by = c("id"))


ggplot(check_corr_titles2, aes(x = factor(project_is_approved), y = score)) + geom_boxplot()

essay1_c <- ggplot(check_corr_titles2,aes(x = score, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill")

# Is there a relationship between the most common value and whether it gets approved?

check_corr_titles <- aggregate(janeaustensentiment$score, by = list(janeaustensentiment$id), FUN = median)
colnames(check_corr_titles) <- c("id", "score")

check_corr_titles2 <- left_join(check_corr_titles, train_text[,1:2], by = c("id"))


ggplot(check_corr_titles2, aes(x = factor(project_is_approved), y = score)) + geom_boxplot()

essay1_d <- ggplot(check_corr_titles2,aes(x = score, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill")
```


# Data exploration on essay 2

## Make data into a tibble and find out how many words are in essay 2

```{r}
# First, we want to select project id, essay 2, and if the project was approved or not

id_title <- c(1, 16, 11)

train_text <- train[,id_title]
train_text[,1] <- as.character(train_text[,1])
train_text[,2] <- as.numeric(train_text[,2])
train_text[,3] <- as.character(train_text[,3])

train_text <- as.tibble(train_text)

tidy_books <- train_text %>% unnest_tokens(word, project_essay_2)

# Find how many words in essay1

freq_table <- count(tidy_books, id) 

title_word_count_by_project <- left_join(freq_table, train_text[,1:2], by = c("id"))

ggplot(title_word_count_by_project, aes(x = factor(project_is_approved), y = n)) + geom_boxplot()

ggplot(title_word_count_by_project, aes(x = n, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill")
```

## Take out common words ("stop words")

```{r}
# Take out stop (common) words

tidy_books <- tidy_books %>%
  anti_join(stop_words)

freq_table <- count(tidy_books, id) 

title_word_count_by_project <- left_join(freq_table, train_text[,1:2], by = c("id"))

ggplot(title_word_count_by_project, aes(x = factor(project_is_approved), y = n)) + geom_boxplot()

essay2_n <- ggplot(title_word_count_by_project, aes(x = n, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill")

# Get a sense of which words are the most common

tidy_books %>%
  count(word, sort = TRUE) 


tidy_books %>%
  count(word, sort = TRUE) %>%
  filter(n > 20000) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip()

```

## Sentiment analysis

### AFINN

```{r}
janeaustensentiment <- tidy_books %>% inner_join(get_sentiments("afinn"))

summary(janeaustensentiment)

janeaustensentiment[1:10,]

# Is there a relationship between the average score and whether or not it gets accepted?

check_corr_titles <- aggregate(janeaustensentiment$score, by = list(janeaustensentiment$id), FUN = mean)
colnames(check_corr_titles) <- c("id", "score")

check_corr_titles2 <- left_join(check_corr_titles, train_text[,1:2], by = c("id"))


ggplot(check_corr_titles2, aes(x = factor(project_is_approved), y = score)) + geom_boxplot()

essay2_a <- ggplot(check_corr_titles2,aes(x = score, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill")

# Is there a relationship between the most positive word and whether it gets approved?

check_corr_titles <- aggregate(janeaustensentiment$score, by = list(janeaustensentiment$id), FUN = max)
colnames(check_corr_titles) <- c("id", "score")

check_corr_titles2 <- left_join(check_corr_titles, train_text[,1:2], by = c("id"))


ggplot(check_corr_titles2, aes(x = factor(project_is_approved), y = score)) + geom_boxplot()

essay2_b <-ggplot(check_corr_titles2,aes(x = score, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill")

# Is there a relationship between the most negative word and whether it gets approved?

check_corr_titles <- aggregate(janeaustensentiment$score, by = list(janeaustensentiment$id), FUN = min)
colnames(check_corr_titles) <- c("id", "score")

check_corr_titles2 <- left_join(check_corr_titles, train_text[,1:2], by = c("id"))


ggplot(check_corr_titles2, aes(x = factor(project_is_approved), y = score)) + geom_boxplot()

essay2_c <-ggplot(check_corr_titles2,aes(x = score, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill")

# Is there a relationship between the most common value and whether it gets approved?

check_corr_titles <- aggregate(janeaustensentiment$score, by = list(janeaustensentiment$id), FUN = median)
colnames(check_corr_titles) <- c("id", "score")

check_corr_titles2 <- left_join(check_corr_titles, train_text[,1:2], by = c("id"))


ggplot(check_corr_titles2, aes(x = factor(project_is_approved), y = score)) + geom_boxplot()

essay2_d <- ggplot(check_corr_titles2,aes(x = score, fill = factor(project_is_approved))) + 
    geom_bar(position = "fill")
```

# Summary Plots

```{r}
plot_grid(title_n, essay1_n, essay2_n, labels = c("A", "B", "C"), ncol = 1)
plot_grid(title_a, essay1_a, essay2_a, labels = c("A", "B", "C"), ncol = 1)
plot_grid(title_b, essay1_b, essay2_b, labels = c("A", "B", "C"), ncol = 1)
plot_grid(title_c, essay1_c, essay2_c, labels = c("A", "B", "C"), ncol = 1)
plot_grid(title_d, essay1_d, essay2_d, labels = c("A", "B", "C"), ncol = 1)
```





## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
